---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# ecoli_regulator_screen_binsfeld25


The goal this repository is to model CPI scores from [Binsfeld et al., 2025](https://www.biorxiv.org/content/10.1101/2024.08.26.609649v1) as a function of regulator presence/absence and compound concentration. This repository contains the workflow followed for noramlization, modeling and cross validation. This repository is found at [rolayoalarcon/ecoli_regulator_screen_binsfeld25](https://github.com/rolayoalarcon/ecoli_regulator_screen_binsfeld25).

## Workflow. 
  
The `workflow` directory contains all Rmd files for this analysis. The complete workflow can be recreated with the following command:  
 
```{r, eval=FALSE}

projthis::proj_workflow_render("workflow")

```

See the description in this directory for more details. 
  
## CPI scores. 
  
The modeled CPI scores can be found in the `CPI_scores` directory. 
